---
title: "spread of organic contamination from a river to a aquifer"
subtitle: "reaction transport model of different compartements in a aquifer"
author: "Stijn Eggenkamp, Niek Geelen, Jarno Herrmann and Jesper van Huizen"
date: "23-3-2022"
output:
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

{r child = ’intro.Rmd’}


# R implementation




# R implementation

Load the required packages.

```{r, message = FALSE}
require(marelac)
require(ReacTran)
```

## The model grid and associated properties

Define the model grid and model parameters that vary in space.

```{r}
# units: time=days, space=meters, amount=moles, concentration=mol/m3

# spatial domain
Length    <- 500    # [m]
N         <- 250    # number of grid cells

# grid with equally sized boxes (not used here, see next R-command)
Grid      <- setup.grid.1D(L = Length, N = N)

# initial conditions
DOM.ini   <- rep(0, length = N) 
NO3.ini   <- rep(0, length = N)
NH3.ini   <- rep(0, length = N)
O2.ini    <- rep(0, length = N)
N2.ini    <- rep(0, length = N)
state.ini <- c(DOM.ini, NO3.ini, NH3.ini, O2.ini, N2.ini)
SVnames   <- c("DOM", "NO3", "NH3", "O2", "N2")
nspec     <- length(SVnames)
```

##Model Parameters
The model parameters for the model
```{r}
parms <- c(
 v        = 10*(24/100),   # [m/d]         the velocity of groundwater
 a        = 1.5,           # [m]           the dispersivity of the aquifer 
 por      = 0.4,           # [-]           average porosity of the aquifer
 NC       = 16/106,        # [-]           Redfield ratio C -> N
 rA       = 0.002*24,      # [/d]          the aerobic respiration rate constant
 kO2      = 20/1000,       # [mol/m3]      affinity constant for O2 substrate
 rD       = 0.002*24,      # [/d]          the denitrification rate constant
 kNO3     = 35/1000,       # [mol/m3]      rate limitation for No3
 rNitri   = 0.36*24,       # [/(mol/m3)/d] nitrification rate constant
 rAer     = 0.0003*24,     # [/d]          the aeration constant
 DOMriver = 0.5,           # [mol/m3]      the concentration of DOM in the river
 O2river  = 210/1000,      # [mol/m3]      the concentration of the oxygen in the river
 NO3river = 100/1000,      # [mol/m3]      the concentration of NO3 in the river
 NH3river = 0/1000,        # [mol/m3]      the concentration of NH3 in the river
 N2up     = 0/1000         # [mol/m3]      the concentration of N2 in the air
)


# calculate the oxygen solubility

O2satmm <- gas_solubility(S = 0, t = 10, species = "O2")      #mmol/m3/bar
O2sat <- O2satmm/1000*0.21                                         #mol/m3/bar
# partial pressure. bar is gehele atmospheric pressure. *21% is only the o2 in de atmosfeer

```


## Definition of the model function

Define the model function that calculates the *time-derivatives* for each state variable. This includes the implementation of the transport and reaction terms.

```{r}
Aquifer <- function (t, state, pars)   # state is a LONG vector
{
  with (as.list(pars),{ 
    
    # unpack state variables
    DOM <- state[ (0*N+1) : (1*N) ]    # first N elements: DOM
    NO3 <- state[ (1*N+1) : (2*N) ]    # next N elements:  NO3
    NH3 <- state[ (2*N+1) : (3*N) ]    # next N elements: NH3
    O2  <- state[ (3*N+1) : (4*N) ]    # next N elements: O2
    N2  <- state[ (4*N+1) : (5*N) ]    # next N elements: N2
    
    # calculate the diffusion coefficient
    disp <- a*v
    
    # === transport rates ===
    # note: zero gradient by default at lower boundaries
    # dissolved substances, VF = porosity    
    tran.DOM <- tran.1D(C = DOM, C.up = DOMriver,    # upper boundary: concentration 
                        dx = Grid, VF = por,         # grid and porosity
                        D = disp, v = v)             # diffusion and advection

    
    tran.NO3 <- tran.1D(C = NO3, C.up = NO3river,    # upper boundary: concentration
                        dx = Grid, VF = por,         # grid and porosity
                        D = disp, v = v)             # diffusive mixing and advection
    
    tran.NH3 <- tran.1D(C = NH3, C.up = NH3river,    # upper boundary: concentration
                        dx = Grid, VF = por,         # grid and porosity
                        D = disp, v = v)             # diffusive mixing and advection
    
    tran.O2  <- tran.1D(C = O2, C.up = O2river,      # upper boundary: concentration
                        dx = Grid, VF = por,         # grid and porosity
                        D = disp, v = v)             # diffusive mixing and advection
    
    tran.N2  <- tran.1D(C = N2, C.up = N2up, 
                        dx = Grid, VF = por, 
                        D = disp, v = v)
    
                                                     
 
    # === reaction rates ===
    # Aerobic Mineralization
    Mineralisation <- rA * DOM * (O2/(O2+kO2))   # [mol DOM/m3 liquid/d] 
    
    # Denitrification
    Denitrification <- rD * DOM * (kO2/(O2+kO2)) * (NO3/(NO3+kNO3)) # [mol DOM/m3 liquid/d]

    
    # Nitrification 
    Nitrification  <- rNitri * NH3 * O2          # [mol/m3 liquid/d]

    # Aeration
    Aeration <- rAer*(O2sat - O2)

    # === mass balances : dC/dt = transport + reactions ===
    
    # dissolved substances
    dDOM.dt  <- tran.DOM$dC - Mineralisation - Denitrification                                     # [molDOM/m3 liquid/d]
    dNO3.dt  <- tran.NO3$dC + Nitrification - (4/5) * Denitrification                              # [molNO3/m3 liquid/d]
    dNH3.dt  <- tran.NH3$dC + NC*Mineralisation + NC*Denitrification - Nitrification               # [molNH3/m3 liquid/d]
    dO2.dt   <- tran.O2$dC - Mineralisation - 2*Nitrification + Aeration                           # [molO2/m3 liquid/d]
    dN2.dt   <- tran.N2$dC + (2/5)*Denitrification
    

    # depth-integrated rates: [mol POC/m2 BULK/d] !!
    TotalMin  <- sum(Mineralisation*Grid$dx*por)
    TotalDen  <- sum(Denitrification*Grid$dx*por)
    TotalNit  <- sum(Nitrification*Grid$dx*por)
    TotalAer  <- sum(Aeration*Grid$dx*por)
  
    return(list(c(dDOM.dt, dNO3.dt, dNH3.dt, dO2.dt, dN2.dt),            # the time-derivatives, as a long vector
          Mineralisation  = Mineralisation,                              # Mineralisation rates, as a vector
          Denitrification = Denitrification,                             # Denitrification rates, as a vector
          Nitrification   = Nitrification,                               # Nitrification rates, as a vector
          Aeration        = Aeration,                                    # Aeration rates, as a vector
          # for creating budgets
          TotalMin      = TotalMin,             # [molDOM/m2 BULK/d]
          TotalDen      = TotalDen,             # [molDOM/m2 BULK/d]
          TotalNit      = TotalNit,             # [mol/m2 BULK/d]
          TotalAer      = TotalAer,
          DOM.SWI.Flux  = tran.DOM$flux.up,     # [mol/m2 BULK/d]
          DOM.Deep.Flux = tran.DOM$flux.down,   # [mol/m2 BULK/d]
          NO3.SWI.Flux  = tran.NO3$flux.up,     # [mol/m2 BULK/d]
          NO3.Deep.Flux = tran.NO3$flux.down,   # [mol/m2 BULK/d]
          NH3.SWI.Flux  = tran.NH3$flux.up,     # [mol/m2 BULK/d]
          NH3.Deep.Flux = tran.NH3$flux.down,   # [mol/m2 BULK/d]
          O2.SWI.Flux   = tran.O2$flux.up,      # [mol/m2 BULK/d]
          O2.Deep.Flux  = tran.O2$flux.down,    # [mol/m2 BULK/d]
          N2.SWI.Flux   = tran.N2$flux.up,      # [mol/m2 BULK/d]
          N2.Deep.Flux  = tran.N2$flux.down))   # [mol/m2 BULK/d]
 })
}
```

## Steady-state solution

Find a steady-state solution with the function *steady.1D* from the package *rootSolve*.

```{r}



std <- steady.1D (y=state.ini, func=Aquifer, parms=parms, 
                   nspec=nspec, dimens=N, names=SVnames,
                   positive = TRUE)     # to have only positive values! 


```

## Plotting

Visualise the steady state solutions, including the depth profile of the process rate.

```{r}
plot(std, grid=Grid$x.mid, lty=1, lwd=2, mfrow=c(2,3), xlab = "length (m)", ylab = "concentration (mol/m3)")
plot(std, grid=Grid$x.mid, which=c("Mineralisation", "Denitrification", "Nitrification", "Aeration"), lty=1, lwd=2, mfrow=c(2,3))


```

## Budget

Check the fluxes at the domain boundaries (only done for the steady-state solution *std1*).

```{r}
toselect <- c("TotalMin", "TotalDen", "TotalNit", "TotalAer", "DOM.SWI.Flux", "DOM.Deep.Flux", 
              "NO3.SWI.Flux", "NO3.Deep.Flux", "NH3.SWI.Flux", "NH3.Deep.Flux", "O2.SWI.Flux", "O2.Deep.Flux",
              "N2.SWI.Flux", "N2.Deep.Flux")
BUDGET   <- std[toselect]
unlist(BUDGET)    # show result
```

Construct the total C budget in the modeled spatial domain to verify that the mass balances are correct: for each state variable, the *difference between fluxes* at the lower and upper boundaries should be **equal** to the *net production rate* integrated along the domain.

```{r}
# DOM mass balance check
(DOM_MBC <- c(flux_difference  =  BUDGET$DOM.SWI.Flux - BUDGET$DOM.Deep.Flux, 
              integrated_rate  = -BUDGET$TotalMin - BUDGET$TotalDen, 
              MBC              =  BUDGET$DOM.SWI.Flux - BUDGET$DOM.Deep.Flux - BUDGET$TotalMin - BUDGET$TotalDen))             # net production


(NO3_MBC <- c(flux_difference  =  BUDGET$NO3.SWI.Flux - BUDGET$NO3.Deep.Flux,
              integrated_rate  =  BUDGET$TotalNit - (4/5) * BUDGET$TotalDen,
              MBC              =  BUDGET$NO3.SWI.Flux - BUDGET$NO3.Deep.Flux + BUDGET$TotalNit - (4/5) * BUDGET$TotalDen))

(NH3_MBC <- c(flux_difference  =  BUDGET$NH3.SWI.Flux - BUDGET$NH3.Deep.Flux,
              integrated_rate  =  (16/106) * BUDGET$TotalDen + (16/106) * BUDGET$TotalMin - BUDGET$TotalNit,
              MBC              =  BUDGET$NH3.SWI.Flux - BUDGET$NH3.Deep.Flux + (16/106) * BUDGET$TotalDen + (16/106) * BUDGET$TotalMin - BUDGET$TotalNit))

(O2_MBC  <- c(flux_difference  =  BUDGET$O2.SWI.Flux - BUDGET$O2.Deep.Flux,
              integrated_rate  = -BUDGET$TotalMin -2*BUDGET$TotalNit + BUDGET$TotalAer,
              MBC              =  BUDGET$O2.SWI.Flux - BUDGET$O2.Deep.Flux + -BUDGET$TotalMin -2*BUDGET$TotalNit + BUDGET$TotalAer))

(N2_MBC  <- c(flux_difference  =  BUDGET$N2.SWI.Flux - BUDGET$N2.Deep.Flux, 
              integrated_rate  = (2/5)*BUDGET$TotalDen, 
              MBC              =  BUDGET$N2.SWI.Flux - BUDGET$N2.Deep.Flux + (2/5)*BUDGET$TotalDen))



```
## The data 
```{r}
data.ext <- read.table(file = "p4-data.dat", header = TRUE, sep = ",")
data.ext$DOM <- data.ext$DOM / parms[["NC"]]           #from N to C

klooiparms <- parms
klooiparms["rNitri"]     <- 0.36*24/2.3
klooiparms["rAer"]       <- 0.0003*24/3.3   
klooiparms["DOMriver"]   <- 0.5*0.70          



std2 <- steady.1D (y=state.ini, func=Aquifer, parms=klooiparms, 
                   nspec=nspec, dimens=N, names=SVnames,
                   positive = TRUE) 

```
```{r}
plot(std, std2, grid=Grid$x.mid, lty=1, lwd=2, mfrow=c(2,3), xlab = "length (m)", ylab = "Concentration (mol/m3)", which = c("DOM","NH3","O2"), obs = data.ext)
plot.new()
legend("right", legend = c("Standard", "Calibration"), lty=1,lwd=2,col=1:3)
```
```{r}
```

